---
title: "Installed Wind Capacity vs. Wind Generation"
output:
  html_document:
    df_print: paged
---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(
 fig.width = 6, fig.height = 6
)
```

Going to use this notebook to look at how installed wind capacity looks as a metric, compared to just straight up wind generation.

```{r clean up, include=FALSE}

rm(list=ls())

```

```{r import, include=FALSE}
library(readxl)
library(readr)
library(ggplot2)
library(ggpubr)
library(lubridate)
library(GGally)
library(knitr)
library(cowplot)
```

```{r load data, include=FALSE}
wind_capacity <- read_excel("data/Capacity_Changes_by_Fuel_Type_Charts_May_2021.xlsx", 
    range = "A74:B97", col_names = c("year", "windcap"))

wind_capacity$windcap <- as.numeric(wind_capacity$windcap)

ercot_dataset <- read.csv("data/ERCOT_Dataset_8June2021.csv", na.strings = ".")
wind_generation <- data.frame(cbind(year = ercot_dataset$year, month = ercot_dataset$month, 
                                    day = ercot_dataset$day, windgen = ercot_dataset$wind))

wind_gen_annual_sum <- aggregate(windgen ~ year,       
                        wind_generation,
                        FUN = sum)

wind_gen_annual_hourly_mean <- aggregate(windgen ~ year,       
                        wind_generation,
                        FUN = mean)

wind_data <- merge(wind_capacity, wind_gen_annual_sum, by = 'year')


wind_data$windcap <- as.numeric(wind_data$windcap)

hourly_windgen_timeseries <- cbind.data.frame(datetime = make_datetime(year=ercot_dataset$year, month =  ercot_dataset$month, day = ercot_dataset$day, hour = ercot_dataset$hourstart), windgen = ercot_dataset$wind)

wind_data$last_years_growth <- as.numeric(diff(wind_capacity$windcap[10:20]))
```

## Data

All data straight from ERCOT - capacity data from their annual capacity changes by fuel type chart, and generation summed annually from their fuel mix report.

```{r echo=FALSE}
wind_data
summary(wind_data[2:4])
```

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=10, fig.width=8}

p <- ggplot(hourly_windgen_timeseries, aes(x=datetime, y=windgen)) +
  geom_point(alpha=0.2, color = "gray") +
  geom_smooth(color = "red") + 
  #geom_bar() +
  xlab("") +
  ylab("Wind Generation (MWh)") +
  ggtitle("Hourlyas Wind Generation in ERCOT", subtitle = "Jan 2011 - Dec 2020. Smoothed with a generalized additive model for vizualization. Data: ERCOT")


p2 <- ggplot(wind_capacity[12:21,], aes(x = year, y=windcap)) +
  #geom_point() + 
  geom_col() +
  #geom_segment(aes(x=year, xend=year, y=0, yend=windcap)) +
  xlab("") +
  ylab("Wind Generation Nameplate Capacity (MW)") +
  ggtitle("Cumulative Annual Installed Wind Generation Capacity in ERCOT", subtitle = "2011 - 2020. Data: ERCOT") +
  scale_x_continuous(breaks=2011:2020,labels=2011:2020)


p3 <- ggplot(wind_data, aes(x = year, y = last_years_growth)) +
  #geom_point() + 
  geom_col() +
  #geom_segment(aes(x=year, xend=year, y=0, yend=windcap)) +
  xlab("") +
  ylab("Wind Generation Nameplate Capacity (MW)") +
  ggtitle("Annual Installation of Wind Generation Capacity in ERCOT", subtitle = "2011 - 2020. Data: ERCOT") +
  scale_x_continuous(breaks=2010:2019,labels=2010:2019)


p4 <- ggpairs(wind_data, columns = c(2:4),lower=list(continuous="smooth"),diag=list(continuous="densityDiag"))

plot_grid(p, p2, p3, p4, labels = NULL)

p4
```


## Testing for Correlation
```{r graphs, echo=FALSE, fig.height=5, fig.width=10, message=FALSE, warning=FALSE}
p5 <- ggscatter(wind_data, x = "last_years_growth", y = "windgen", 
                add = "reg.line", conf.int = TRUE, 
                cor.coef = TRUE, cor.method = "pearson",
                xlab = "Prior Year's growth in windgen Nameplate Capacity (MW) from", 
                ylab = "Total Annual Wind Generation in ERCOT (MWh)") +
                theme_bw() +
                ggtitle("Correlating Wind Generation and Wind Capacity in ERCOT", 
                        subtitle = "2011 - 2020. Data: ERCOT")


p6 <- ggscatter(wind_data, x = "windcap", y = "windgen", 
                add = "reg.line", conf.int = TRUE, 
                cor.coef = TRUE, cor.method = "pearson",
                xlab = "Cumulative Installed Wind Generation Capacity in ERCOT (MW)", 
                ylab = "Total Annual Wind Generation in ERCOT (MWh)") +
                theme_bw() +
                ggtitle("Correlating Wind Generation and Wind Capacity in ERCOT", 
                        subtitle = "2011 - 2020. Data: ERCOT")

plot_grid(p5, p6, labels = NULL)


```

### Normalcy

Let's see if this data is normally distributed before testing for correlation.

```{r echo = FALSE, fig.width=12, fig.height=6}
p7 <- ggplot(wind_data, aes(sample = windcap)) +
      stat_qq() + 
      stat_qq_line() +
      ggtitle("QQ Plot for Wind Capacity")

p8 <- ggplot(wind_data, aes(sample = windgen)) +
      stat_qq() + 
      stat_qq_line() +
      ggtitle("QQ Plot for Wind Generation")

p9 <- ggplot(wind_data, aes(sample = last_years_growth)) +
      stat_qq() + 
      stat_qq_line() +
      ggtitle("QQ Plot for Annual Wind Capacity Additions")

plot_grid(p7, p8, p9, nrow=1, ncol=3, rel_widths = 5,labels = NULL)

```


```{r echo = FALSE}
# Shapiro-Wilk normality test for wind capacity:
shapiro.test(wind_data$windcap) # => p = 0.3536
# Shapiro-Wilk normality test for summed generation:
shapiro.test(wind_data$windgen) # => p = 0.2721
# Shapiro-Wilk normality test for mean hourly generation:
shapiro.test(wind_data$last_years_growth) # => p = 0.4353
```

### Pearson Test

Normally distributed, so we can test for correlation with Pearson.

```{r echo = FALSE}
res1 <- cor.test(wind_data$windcap, wind_data$windgen, 
                    method = "pearson")
res1

res2 <- cor.test(wind_data$windgen, wind_data$last_years_growth, 
                    method = "pearson")
res2

```

## Conclusion

With a p-value of 1.4e-07 and a correlation coefficient of 0.986485, we can conclude generation and capacity are significantly  linearly correlated.

However, with a p-value of 0.06545 and a correlation coefficient of 0.6021706, annual capacity additions do not significantly linearly correlate with wind generation in ERCOT.


